---
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---
## SETUP
```{r Chunk Options}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r Load Packages}
library(tidyverse)
library(tidymodels)
library(readxl)
library(lubridate)
library(extrafont)
library(latex2exp)
library(gridExtra)
library(cowplot)
library(patchwork)
library(grid)
```

```{r Load HW Data}
hw = read_excel("HW-Days.xlsx", col_names = TRUE)
```


## ASHEVILLE
```{r Load AVL Data}
avl <- read_excel("KAVL-heatindex-1971-2021.xlsx", col_names = TRUE)
```

```{r AVL Dataframe Cols}
# convert datetime col to datetime object
avl <- avl |>
  mutate(datetime = as.Date(datetime))

# create month-day column for grouping
avl$month_day <- format(avl$datetime, format = "%m-%d") # month-day col

avl$month <- as.numeric(format(avl$datetime, format = "%m"))

# create year column
avl$year <- format(avl$datetime, format = "%Y")

# drop nas
avl <- drop_na(avl)
```

```{r filter AVL for summer months}
# filter for summer months
avl_summer = avl |>
  filter(
    month > 2 & month < 11
  )
```

```{r AVL 98P heat index max and min}
# create cols storing percentile values
avl = avl |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r AVL 98P heat index max and min summer}
# summer months
avl_summer = avl_summer |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r AVL add > 98p indicators to df}
# new cols indicating extreme heat
avl = avl |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r AVL summer add extreme indicators to df}
# summer months
avl_summer = avl_summer |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r AVL add extreme heat day indicators to df max heat index}
avl <- avl |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r AVL add extreme heat day indicators to df min heat index}
avl <- avl |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r AVL add extreme heat day indicators to df max heat index summer}
avl_summer = avl_summer |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r AVL add extreme heat day indicators to df min heat index summer}
avl_summer = avl_summer |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r AVL Summarize Extreme Day Counts}
avl_extreme_counts = avl |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

```{r AVL Summarize Extreme Day Counts Summer}
avl_summer_extreme_counts = avl_summer |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

### Plots: Extreme Days (Heat Index)
```{r AVL Max Days Bar Graph}
# plot extreme days each year
# bar plot

ggplot(avl_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Asheville",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r AVL Summer Max Days Bar graph}
# extreme days in summer months, by daily max heat index 

ggplot(avl_summer_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Asheville",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r AVL Min Days Bar Graph}
ggplot(avl_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Asheville", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r AVL Summer Min Days Bar Graph}
ggplot(avl_summer_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Asheville",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting AVL Extremes to plot}
avl_extreme_pivot = avl_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r AVL Extreme Days Min and Max Bar Graph}
avl_extreme_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Asheville, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r AVL Extreme Min and Max Loess and Points}
lo.avl.max = lowess(avl_extreme_counts$count_extreme_max ~ avl_extreme_counts$year, f = 1/5)
lo.avl.min = lowess(avl_extreme_counts$count_extreme_min ~ avl_extreme_counts$year, f = 1/5)

# Create a ggplot
ggplot(avl_extreme_counts, aes(x = year, y = count_extreme_max)) +
  geom_point(color = "red", size = 1, alpha = 0.3) + 
  geom_point(aes(x = year, y = count_extreme_min), color = "blue", size = 1, alpha = 0.3) + 
  geom_line(aes(y = lo.avl.max$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.avl.min$y, group = 1), color = "blue", linewidth = 0.5) +
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Asheville"
    ) + 
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 30, by = 5), limits = c(0, 30)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting AVL Summer Extremes to plot}
avl_summer_pivot = avl_summer_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r AVL Summer Extreme Days Min and Max Bar Graph}
avl_summer_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Asheville, Mar-Oct, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r AVL Summer Extreme Min and Max Loess and Points}
lo.avl.max.sum = lowess(avl_summer_extreme_counts$count_extreme_max ~ avl_summer_extreme_counts$year, f = 1/5)
lo.avl.min.sum = lowess(avl_summer_extreme_counts$count_extreme_min ~ avl_summer_extreme_counts$year, f = 1/5)

ggplot(avl_summer_extreme_counts, aes(x = year, y = count_extreme_max)) +
  geom_point(color = "red", alpha = 0.5, size = 1) + 
  geom_line(aes(y = lo.avl.max.sum$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.avl.min.sum$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Extreme Heat Days", 
    subtitle = "Asheville, Mar-Oct, 1971-2021") + 
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_minimal(base_family = "Times")
``` 

### Plots: HW Days
```{r AVL HW Days by Max HI Bar Graph}
avl_hw_max_plot = hw |>
ggplot(aes(x = year, y = count_avl_max)) +
  geom_col(fill = "red") +
  labs(
    title = "Asheville",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r AVL HW Days by Min HI Bar Graph}
avl_hw_min_plot = hw |>
ggplot(aes(x = year, y = count_avl_min)) +
  geom_col(fill = "blue") +
  labs(
    title = "Asheville",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting AVL HW Days to Plot}
avl_hw_pivot = hw |>
  select(c(year, count_avl_max, count_avl_min)) |>
  pivot_longer(
    cols = c(count_avl_max, count_avl_min), 
    names_to = "Measurement", 
    values_to = "Count"
  )
```

```{r AVL HW Days by Min AND Max Bar Graph}
avl_hw_pivot |>
  ggplot(aes(x = year, y = Count, fill = Measurement)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Heatwave Days", 
    subtitle = "Asheville, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_avl_max", "count_avl_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

## CHARLOTTE
```{r Load CLT Data}
clt <- read_excel("KLCT-heatindex-1971-2021.xlsx", col_names = TRUE)
```

```{r CLT Dataframe Cols}
# convert datetime col to datetime object
clt <- clt |>
  mutate(datetime = as.Date(datetime))

# create month-day column for grouping
clt$month_day <- format(clt$datetime, format = "%m-%d") # month-day col

clt$month <- as.numeric(format(clt$datetime, format = "%m"))

# create year column
clt$year <- format(clt$datetime, format = "%Y")

# numeric
clt$heatindexmax2m = as.numeric(clt$heatindexmax2m)
clt$heatindexmin2m = as.numeric(clt$heatindexmin2m)

# drop nas
clt <- drop_na(clt)
```

```{r filter for CLT summer months}
# filter for summer months
clt_summer = clt |>
  filter(
    month > 2 & month < 11
  )
```

```{r CLT 98P heat index max and min}
# create cols storing percentile values
clt <- clt |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r CLT 98P heat index max and min summer}
# summer months
clt_summer = clt_summer |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r CLT add > 98P indicators}
# new cols indicating extreme heat
clt <- clt |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r CLT summer add > 98P indicators}
# summer months
clt_summer = clt_summer |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r CLT add extreme heat day indicators to df max hi}
clt <- clt |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r CLT add extreme heat day indicators to df min hi}
clt <- clt |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r CLT add extreme heat day indicators to df max hi summer}
clt_summer = clt_summer |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r CLT add extreme heat day indicators to df min hi summer}
clt_summer = clt_summer |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r CLT Summarize Extreme Day Counts}
clt_extreme_counts = clt |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

```{r CLT Summarize Extreme Day Counts Summer}
clt_summer_extreme_counts = clt_summer |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

### Plots: Extremes Heat Index
```{r CLT Max Days Bar Graph}
# plot extreme days each year
# bar plot

clt_hi_max_plot = ggplot(clt_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Charlotte", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +  
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r CLT Summer Max Days Bar Graph}
# extreme days in summer months, by daily max heat index 
ggplot(clt_summer_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Charlotte",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r CLT Min Days Bar Graph}
clt_hi_min_plot = ggplot(clt_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Charlotte", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r CLT Summer Min Days Bar Graph}
ggplot(clt_summer_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Charlotte", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +  
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting CLT Extremes to plot}
clt_extreme_pivot = clt_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r CLT Extreme Days Min and Max Bar Graph}
clt_extreme_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Charlotte, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r CLT Extreme Min and Max Loess and Points}
lo.clt.max = lowess(clt_extreme_counts$count_extreme_max ~ clt_extreme_counts$year, f = 1/5)
lo.clt.min = lowess(clt_extreme_counts$count_extreme_min ~ clt_extreme_counts$year, f = 1/5)

# Create a ggplot
ggplot(clt_extreme_counts, aes(x = year, y = count_extreme_max))  +
  geom_point(color = "red", size = 1, alpha = 0.3) + 
  geom_point(aes(x = year, y = count_extreme_min), color = "blue", size = 1, alpha = 0.3) + 
  geom_line(aes(y = lo.clt.max$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.clt.min$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Charlotte") + 
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 30, by = 5), limits = c(0, 30)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting CLT Summer Extremes to Plot}
clt_summer_pivot = clt_summer_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r CLT Summer Extreme Days Min and Max Bar Graph}
clt_summer_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Charlotte, Mar-Oct, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r CLT Summer Extreme Min and Max Loess and Points}
lo.clt.max.sum = lowess(clt_summer_extreme_counts$count_extreme_max ~ clt_summer_extreme_counts$year, f = 1/5)
lo.clt.min.sum = lowess(clt_summer_extreme_counts$count_extreme_min ~ clt_summer_extreme_counts$year, f = 1/5)

ggplot(clt_summer_extreme_counts, aes(x = year, y = count_extreme_max)) +
  geom_line(aes(y = lo.clt.max.sum$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.clt.min.sum$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Extreme Heat Days", 
    subtitle = "Charlotte, Mar-Oct, 1971-2021") + 
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_minimal(base_family = "Times")
``` 

### Plots: HW Days
```{r CLT HW Days by Max HI Bar Graph}
clt_hw_max_plot = hw |>
ggplot(aes(x = year, y = count_clt_max)) +
  geom_col(fill = "red") +
  labs(
    title = "Charlotte", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r CLT HW Days by Min HI Bar Graph}
clt_hw_min_plot = hw |>
ggplot(aes(x = year, y = count_clt_min)) +
  geom_col(fill = "blue") +
  labs(
    title = "Charlotte", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting CLT HW Days to Plot}
clt_hw_pivot = hw |>
  select(c(year, count_clt_max, count_clt_min)) |>
  pivot_longer(
    cols = c(count_clt_max, count_clt_min), 
    names_to = "Measurement", 
    values_to = "Count"
  )
```

```{r CLT HW Days by Min AND Max Bar Graph}
clt_hw_pivot |>
  ggplot(aes(x = year, y = Count, fill = Measurement)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Heatwave Days", 
    subtitle = "Charlotte, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_clt_max", "count_clt_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

## GREENSBORO
```{r Load GSO Data}
gso <- read_excel("KGSO-heatindex-1971-2021.xlsx", col_names = TRUE)
```

```{r GSO Dataframe Cols}
# convert datetime col to datetime object
gso <- gso |>
  mutate(datetime = as.Date(datetime))

# create month-day column for grouping
gso$month_day <- format(gso$datetime, format = "%m-%d") # month-day col

gso$month <- as.numeric(format(gso$datetime, format = "%m"))

# create year column
gso$year <- format(gso$datetime, format = "%Y")

# numeric
gso$heatindexmax2m = as.numeric(gso$heatindexmax2m)
gso$heatindexmin2m = as.numeric(gso$heatindexmin2m)

# drop nas
gso <- drop_na(gso)
```

```{r filter GSO for summer months}
# filter for summer months
gso_summer = gso |>
  filter(
    month > 2 & month < 11
  )
```

```{r GSO 98P calculation heat index max and min}
# create cols storing percentile values
gso <- gso |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r GSO 98P calculation heat index max and min summer}
# summer months
gso_summer = gso_summer |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r GSO add > 98P indicator}
# new cols indicating extreme heat
gso <- gso |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r GSO add > 98P indicator summer}
# summer months
gso_summer = gso_summer |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r GSO add consecutive extreme days indicator max hi}
gso <- gso |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r GSO add consecutive extreme days indicator min hi}
gso <- gso |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r GSO add consecutive extreme days indicator max hi summer}
gso_summer = gso_summer |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r GSO add consecutive extreme days indicator min hi summer}
gso_summer = gso_summer |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r Summarize GSO Extremes}
gso_extreme_counts = gso |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

```{r Summarize GSO Summer Extremes}
gso_summer_extreme_counts = gso_summer |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

### Plots, Extreme HI
```{r GSO Max HI Days Bar Graph}
# plot extreme days each year
# bar plot

gso_hi_max_plot = ggplot(gso_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Greensboro", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r GSO Max HI Days Summer Bar Graph}
# extreme days in summer months, by daily max heat index 
ggplot(gso_summer_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Greensboro",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r GSO Min HI Days Bar Graph}
gso_hi_min_plot = ggplot(gso_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Greensboro", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r GSO Min HI Days Summer Bar Graph}
ggplot(gso_summer_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Greensboro",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting GSO extremes to plot}
gso_extreme_pivot = gso_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r GSO Extreme Days Min and Max Bar Graph}
gso_extreme_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Greensboro, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r GSO Extreme Min and Max Loess and Points}
lo.gso.max = lowess(gso_extreme_counts$count_extreme_max ~ gso_extreme_counts$year, f = 1/5)
lo.gso.min = lowess(gso_extreme_counts$count_extreme_min ~ gso_extreme_counts$year, f = 1/5)

# Create a ggplot
ggplot(gso_extreme_counts, aes(x = year, y = count_extreme_max))  +
  geom_point(color = "red", size = 1, alpha = 0.3) + 
  geom_point(aes(x = year, y = count_extreme_min), color = "blue", size = 1, alpha = 0.3) + 
  geom_line(aes(y = lo.gso.max$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.gso.min$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Greensboro") + 
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 30, by = 5), limits = c(0, 30)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting GSO Summer Extremes to plot}
gso_summer_pivot = gso_summer_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r GSO Summer Extreme Days Min and Max Bar Graph}
gso_summer_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Greensboro, Mar-Oct, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r GSO Summer Extreme Min and Max Loess and Points}
lo.gso.max.sum = lowess(gso_summer_extreme_counts$count_extreme_max ~ gso_summer_extreme_counts$year, f = 1/5)
lo.gso.min.sum = lowess(gso_summer_extreme_counts$count_extreme_min ~ gso_summer_extreme_counts$year, f = 1/5)

ggplot(gso_summer_extreme_counts, aes(x = year, y = count_extreme_max)) +
  geom_line(aes(y = lo.gso.max.sum$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.gso.min.sum$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Extreme Heat Days", 
    subtitle = "Greensboro, Mar-Oct, 1971-2021") + 
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_minimal(base_family = "Times")
``` 

### Plots: HW Days
```{r GSO HW Days by Max HI Bar Graph}
gso_hw_max_plot = hw |>
ggplot(aes(x = year, y = count_gso_max)) +
  geom_col(fill = "red") +
  labs(
    title = "Greensboro",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r GSO HW Days by Min HI Bar Graph}
gso_hw_min_plot = hw |>
ggplot(aes(x = year, y = count_gso_min)) +
  geom_col(fill = "blue") +
  labs(
    title = "Greensboro", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting GSO HW Days to Plot}
gso_hw_pivot = hw |>
  select(c(year, count_gso_max, count_gso_min)) |>
  pivot_longer(
    cols = c(count_gso_max, count_gso_min), 
    names_to = "Measurement", 
    values_to = "Count"
  )
```

```{r GSO HW Days by Min AND Max Bar Graph}
gso_hw_pivot |>
  ggplot(aes(x = year, y = Count, fill = Measurement)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Heatwave Days", 
    subtitle = "Greensboro, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_gso_max", "count_gso_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

## RALEIGH-DURHAM
```{r Load RDU Data}
rdu <- read_excel("KRDU-heat-index-1971-2021.xlsx", col_names = TRUE)
```

```{r RDU Dataframe Cols}
# convert datetime col to datetime object
rdu <- rdu |>
  mutate(datetime = as.Date(date))

# create month-day column for grouping
rdu$month_day <- format(rdu$datetime, format = "%m-%d") # month-day col

rdu$month <- as.numeric(format(rdu$datetime, format = "%m"))

# create year column
rdu$year <- format(rdu$datetime, format = "%Y")

# numeric
rdu$heatindexmax2m = as.numeric(rdu$heatindexmax2m)
rdu$heatindexmin2m = as.numeric(rdu$heatindexmin2m)

# drop nas
rdu <- drop_na(rdu)
```

```{r filter RDU for summer months}
# filter for summer months
rdu_summer = rdu |>
  filter(
    month > 2 & month < 11
  )
```

```{r RDU 98P calculation heat index max and min}
# create cols storing percentile values
rdu <- rdu |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r RDU 98P calculation heat index max and min summer}
# summer months
rdu_summer = rdu_summer |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r RDU add > 98P indicator}
# new cols indicating extreme heat
rdu <- rdu |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r RDU add > 98P indicator summer}
# summer months
rdu_summer = rdu_summer |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r RDU add consecutive extreme days indicator max hi}
rdu <- rdu |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r RDU add consecutive extreme days indicator min hi}
rdu <- rdu |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r RDU add consecutive extreme days indicator max hi summer}
rdu_summer = rdu_summer |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r RDU add consecutive extreme days indicator min hi summer}
rdu_summer = rdu_summer |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r Summarize RDU Extremes}
rdu_extreme_counts = rdu |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

```{r Summarize RDU Summer Extremes}
rdu_summer_extreme_counts = rdu_summer |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

### Plots, Extreme HI
```{r RDU Max HI Days Bar Graph}
# plot extreme days each year
# bar plot
ggplot(rdu_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Raleigh-Durham", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r RDU Max HI Days Summer Bar Graph}
# extreme days in summer months, by daily max heat index 
ggplot(rdu_summer_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Raleigh-Durham",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r RDU Min HI Days Bar Graph}
ggplot(rdu_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Raleigh-Durham", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r RDU Min HI Days Summer Bar Graph}
ggplot(rdu_summer_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Raleigh-Durham",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting RDU extremes to plot}
rdu_extreme_pivot = rdu_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r RDU Extreme Days Min and Max Bar Graph}
rdu_extreme_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Raleigh-Durham, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r RDU Extreme Min and Max Loess and Points}
lo.rdu.max = lowess(rdu_extreme_counts$count_extreme_max ~ rdu_extreme_counts$year, f = 1/5)
lo.rdu.min = lowess(rdu_extreme_counts$count_extreme_min ~ rdu_extreme_counts$year, f = 1/5)

# Create a ggplot
ggplot(rdu_extreme_counts, aes(x = year, y = count_extreme_max))  +
  geom_point(color = "red", size = 1, alpha = 0.3) + 
  geom_point(aes(x = year, y = count_extreme_min), color = "blue", size = 1, alpha = 0.3) + 
  geom_line(aes(y = lo.rdu.max$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.rdu.min$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Raleigh-Durham") + 
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 30, by = 5), limits = c(0, 30)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting RDU Summer Extremes to plot}
rdu_summer_pivot = rdu_summer_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r RDU Summer Extreme Days Min and Max Bar Graph}
gso_summer_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Raleigh-Durham, Mar-Oct, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r RDU Summer Extreme Min and Max Loess and Points}
lo.rdu.max.sum = lowess(rdu_summer_extreme_counts$count_extreme_max ~ rdu_summer_extreme_counts$year, f = 1/5)
lo.rdu.min.sum = lowess(rdu_summer_extreme_counts$count_extreme_min ~ rdu_summer_extreme_counts$year, f = 1/5)

ggplot(rdu_summer_extreme_counts, aes(x = year, y = count_extreme_max)) +
  geom_line(aes(y = lo.rdu.max.sum$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.rdu.min.sum$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Extreme Heat Days", 
    subtitle = "Raleigh-Durham, Mar-Oct, 1971-2021") + 
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_minimal(base_family = "Times")
``` 
### Plots: HW Days
```{r RDU HW Days by Max HI Bar Graph}
rdu_hw_max_plot = hw |>
ggplot(aes(x = year, y = count_rdu_max)) +
  geom_col(fill = "red") +
  labs(
    title = "Raleigh-Durham", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r RDU HW Days by Min HI Bar Graph}
rdu_hw_min_plot = hw |>
ggplot(aes(x = year, y = count_rdu_min)) +
  geom_col(fill = "blue") +
  labs(
    title = "Raleigh-Durham",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting RDU HW Days to Plot}
rdu_hw_pivot = hw |>
  select(c(year, count_rdu_max, count_rdu_min)) |>
  pivot_longer(
    cols = c(count_rdu_max, count_rdu_min), 
    names_to = "Measurement", 
    values_to = "Count"
  )
```

```{r RDU HW Days by Min AND Max Bar graph}
rdu_hw_pivot |>
  ggplot(aes(x = year, y = Count, fill = Measurement)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Heatwave Days", 
    subtitle = "Raleigh-Durham, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_rdu_max", "count_rdu_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

## WILMINGTON
```{r Load WILM Data}
wilm <- read_excel("KILM-heatindex-1971-2021.xlsx", col_names = TRUE)
```

```{r WILM Dataframe cols}
# convert datetime col to datetime object
wilm <- wilm |>
  mutate(datetime = as.Date(datetime))

# create month-day column for grouping
wilm$month_day <- format(wilm$datetime, format = "%m-%d") # month-day col

wilm$month <- as.numeric(format(wilm$datetime, format = "%m"))

# create year column
wilm$year <- format(wilm$datetime, format = "%Y")

# numeric
wilm$heatindexmax2m = as.numeric(wilm$heatindexmax2m)
wilm$heatindexmin2m = as.numeric(wilm$heatindexmin2m)

# drop nas
wilm <- drop_na(wilm)
```

```{r filter Wilm for summer months}
# filter for summer months
wilm_summer = wilm |>
  filter(
    month > 2 & month < 11
  )
```

```{r Wilm calculate 98p heat index min and max}
# create cols storing percentile values
wilm <- wilm |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r Wilm 98p heat index max and min summer}
# summer months
wilm_summer = wilm_summer |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r WILM add > 98p indicators to df}
# new cols indicating extreme heat
wilm <- wilm |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r WILM add > 98p indicators to df summer}
# summer months
wilm_summer = wilm_summer |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r WILM add extreme day indicators to df hi max}
wilm <- wilm |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r WILM add extreme day indicators to df hi min}
wilm <- wilm |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r WIlM add extreme day indicators to df hi max summer}
wilm_summer = wilm_summer |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r WIlM add extreme day indicators to df hi min summer}
wilm_summer = wilm_summer |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r WILM summarize extreme counts}
wilm_extreme_counts = wilm |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

```{r WILM summarize extreme counts summer}
wilm_summer_extreme_counts = wilm_summer |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```


### Plots, Extreme HI
#### Max HI Days
```{r wilm Max Bar}
# plot extreme days each year
# bar plot

wilm_hi_max_plot = ggplot(wilm_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Wilmington", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r WILM summer Max Days bar graph}
# extreme days in summer months, by daily max heat index 
ggplot(wilm_summer_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Wilmington",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r WILM min Days bar graph}
wilm_hi_min_plot = ggplot(wilm_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Wilmington", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) +
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r WILM summer min days bar graph}
ggplot(wilm_summer_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Wilmington", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting WILM Extremes to plot}
wilm_extreme_pivot = wilm_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r WILM Extreme Days Min and Max Bar graph}
wilm_extreme_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Wilmington, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r WILM Extreme Min adn Max Loess and Points}
lo.wilm.max = lowess(wilm_extreme_counts$count_extreme_max ~ wilm_extreme_counts$year, f = 1/5)
lo.wilm.min = lowess(wilm_extreme_counts$count_extreme_min ~ wilm_extreme_counts$year, f = 1/5)

# Create a ggplot
ggplot(wilm_extreme_counts, aes(x = year, y = count_extreme_max)) +
  geom_point(color = "red", size = 1, alpha = 0.3) + 
  geom_point(aes(x = year, y = count_extreme_min), color = "blue", size = 1, alpha = 0.3) + 
  geom_line(aes(y = lo.wilm.max$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.wilm.min$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Wilmington") + 
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 30, by = 5), limits = c(0, 30)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivoting WILM summer Extremes to plot}
wilm_summer_pivot = wilm_summer_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r WILM Summer Extreme Days Min and Max Bar Graph}
wilm_summer_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Wilmington, Mar-Oct, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r WILM Summer Extreme Min and Max Loess and Points}
lo.wilm.max.sum = lowess(wilm_summer_extreme_counts$count_extreme_max ~ wilm_summer_extreme_counts$year, f = 1/5)
lo.wilm.min.sum = lowess(wilm_summer_extreme_counts$count_extreme_min ~ wilm_summer_extreme_counts$year, f = 1/5)

ggplot(wilm_summer_extreme_counts, aes(x = year, y = count_extreme_max)) +
  geom_line(aes(y = lo.gso.max.sum$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.gso.min.sum$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Extreme Heat Days", 
    subtitle = "Wilmington, Mar-Oct, 1971-2021") + 
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_minimal(base_family = "Times")
``` 

### Plots: HW Days
```{r wilm Max HW Bar}
wilm_hw_max_plot = hw |>
ggplot(aes(x = year, y = count_kilm_max)) +
  geom_col(fill = "red") +
  labs(
    title = "Wilmington", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r wilm Min HW Bar}
wilm_hw_min_plot = hw |>
ggplot(aes(x = year, y = count_kilm_min)) +
  geom_col(fill = "blue") +
  labs(
    title = "Wilmington", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r wilm pivoting hw to plot}
wilm_hw_pivot = hw |>
  select(c(year, count_kilm_max, count_kilm_min)) |>
  pivot_longer(
    cols = c(count_kilm_max, count_kilm_min), 
    names_to = "Measurement", 
    values_to = "Count"
  )
```

```{r wilm Min and Max hw bar graph}
wilm_hw_pivot |>
  ggplot(aes(x = year, y = Count, fill = Measurement)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Heatwave Days", 
    subtitle = "Wilmington, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_kilm_max", "count_kilm_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

## HATTERAS
```{r Load hse Data}
hse <- read_excel("KHSE-heatindex-1971-2021.xlsx", col_names = TRUE)
```

```{r hse dataframe cols}
# convert datetime col to datetime object
hse <- hse |>
  mutate(datetime = as.Date(datetime))

# create month-day column for grouping
hse$month_day <- format(hse$datetime, format = "%m-%d") # month-day col

hse$month <- as.numeric(format(hse$datetime, format = "%m"))

# create year column
hse$year <- format(hse$datetime, format = "%Y")

# numeric
hse$heatindexmax2m = as.numeric(hse$heatindexmax2m)
hse$heatindexmin2m = as.numeric(hse$heatindexmin2m)

# drop nas
hse <- drop_na(hse)
```

```{r hse filter for summer}
# filter for summer months
hse_summer = hse |>
  filter(
    month > 2 & month < 11
  )
```

```{r hse Percentile Calc}
# create cols storing percentile values
hse <- hse |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r hse summer percentile calc}
# summer months
hse_summer = hse_summer |>
  group_by(month_day) |>
  mutate(`Max 98th P` = quantile(heatindexmax2m, 0.98), 
         `Min 98th P` = quantile(heatindexmin2m, 0.98)) |>
  ungroup()
```

```{r hse tag days exceeding 98p}
# new cols indicating extreme heat
hse <- hse |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r hse tag summer days exceeding 98p}
# summer months
hse_summer = hse_summer |>
  mutate(
    himax_hw = ifelse(heatindexmax2m > `Max 98th P`, 1, 0), # for max HI
    himin_hw = ifelse(heatindexmin2m > `Min 98th P`, 1, 0) # for min HI
  )
```

```{r hse tag days exceeding 98p 2 consec days by max}
hse <- hse |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r hse tag days exceeding 98p 2 consec days by min}
hse <- hse |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r hse tag days exceeding 98p 2 consec days by max summer}
hse_summer = hse_summer |>
  group_by(ID_MAX = data.table::rleid(himax_hw == 1)) |>
  mutate(
    consec_himax = if_else(
      (himax_hw == 1 & (dplyr::lag(himax_hw, default = first(himax_hw)) == 1)) | 
      (himax_hw == 1 & (dplyr::lead(himax_hw, default = last(himax_hw)) == 1)), 
      1, 0
    )
  )
```

```{r hse tag days exceeding 98p 2 consec days by min summer}
hse_summer = hse_summer |>
  group_by(ID_MIN = data.table::rleid(himin_hw == 1)) |>
  mutate(
    consec_himin = if_else(
      (himin_hw == 1 & (dplyr::lag(himin_hw, default = first(himin_hw)) == 1)) | 
      (himin_hw == 1 & (dplyr::lead(himin_hw, default = last(himin_hw)) == 1)), 
      1, 0
    )
  )
```

```{r hse summarize extremes}
hse_extreme_counts = hse |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```

```{r hse summarzie extremes summer}
hse_summer_extreme_counts = hse_summer |>
  group_by(year) |>
  summarize(
    count_extreme_max = sum(consec_himax > 0), 
    count_extreme_min = sum(consec_himin > 0)
    )
```


### Plots, Extreme HI
```{r hse days max hi > 98p bar graph}
# plot extreme days each year
# bar plot

hse_hi_max_plot = ggplot(hse_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Cape Hatteras", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) + 
  theme_minimal(base_family = "Times") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
# extreme days in summer months, by daily max heat index 
ggplot(hse_summer_extreme_counts, aes(x = factor(year), y = count_extreme_max)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(
    title = "Cape Hatteras", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r hse days min hi > 98p bar graph}
hse_hi_min_plot = ggplot(hse_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Cape Hatteras",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) +
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(hse_summer_extreme_counts, aes(x = factor(year), y = count_extreme_min)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(
    title = "Cape Hatteras", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) +
  scale_y_continuous(breaks = seq(0, 35, by = 5), limits = c(0, 35)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r Pivot hse to plot both max and min}
hse_extreme_pivot = hse_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r hse Min Max Bar}
hse_extreme_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Cape Hatteras, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r hse max min hi loess}
lo.hse.max = lowess(hse_extreme_counts$count_extreme_max ~ hse_extreme_counts$year, f = 1/5)
lo.hse.min = lowess(hse_extreme_counts$count_extreme_min ~ hse_extreme_counts$year, f = 1/5)

# Create a ggplot
ggplot(hse_extreme_counts, aes(x = year, y = count_extreme_max)) +
  geom_point(color = "red", size = 1, alpha = 0.3) + 
  geom_point(aes(x = year, y = count_extreme_min), color = "blue", size = 1, alpha = 0.3) + 
  geom_line(aes(y = lo.hse.max$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.hse.min$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Cape Hatteras") + 
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10))+ 
  scale_y_continuous(breaks = seq(0, 30, by = 5), limits = c(0, 30)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r pivot hse summer to plot both max and min}
hse_summer_pivot = hse_summer_extreme_counts |>
  pivot_longer(
    cols = c(count_extreme_max, count_extreme_min), 
    names_to = "measurement", 
    values_to = "count"
  )
```

```{r hse max AND min hi summer bar graph}
hse_summer_pivot |>
  ggplot(aes(x = factor(year), y = count, fill = measurement)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Days Heat Indices > 98 P", 
    subtitle = "Cape Hatteras, Mar-Oct, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_extreme_max", "count_extreme_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

```{r hse max AND min hi summer loess}
lo.hse.max.sum = lowess(hse_summer_extreme_counts$count_extreme_max ~ hse_summer_extreme_counts$year, f = 1/5)
lo.hse.min.sum = lowess(hse_summer_extreme_counts$count_extreme_min ~ hse_summer_extreme_counts$year, f = 1/5)

ggplot(hse_summer_extreme_counts, aes(x = year, y = count_extreme_max)) +
  geom_line(aes(y = lo.hse.max.sum$y, group = 1), color = "red", linewidth = 0.5) +
  geom_line(aes(y = lo.hse.min.sum$y, group = 1), color = "blue", linewidth = 0.5) + 
  labs(
    x = "Year",
    y = "Number of Days",
    title = "Extreme Heat Days", 
    subtitle = "Cape Hatteras, Mar-Oct, 1971-2021") + 
  theme_minimal(base_family = "Times") +
  scale_x_discrete(breaks = seq(1971, 2021, by = 10)) + 
  scale_color_manual(
    values = c("HI Max" = "red", "HI Min" = "blue"),
    name = "Measurement"
  ) +
  theme_minimal(base_family = "Times")
``` 

### Plots: HW Days
```{r hse hw days by max hi bar graph}
hse_hw_max_plot = hw |>
ggplot(aes(x = year, y = count_hse_max)) +
  geom_col(fill = "red") +
  labs(
    title = "Cape Hatteras", 
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r hse hw days by min hi bar graph}
hse_hw_min_plot = hw |>
ggplot(aes(x = year, y = count_hse_min)) +
  geom_col(fill = "blue") +
  labs(
    title = "Cape Hatteras",
    x = "Year", 
    y = "Number of Days"
  ) +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10))  + 
  scale_y_continuous(breaks = seq(0, 15, by = 5), limits = c(0, 15)) + 
  theme_minimal(base_family = "Times") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r hse pivot hw days}
hse_hw_pivot = hw |>
  select(c(year, count_hse_max, count_hse_min)) |>
  pivot_longer(
    cols = c(count_hse_max, count_hse_min), 
    names_to = "Measurement", 
    values_to = "Count"
  )
```

```{r hse hw days min and max}
hse_hw_pivot |>
  ggplot(aes(x = year, y = Count, fill = Measurement)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "Year", 
    y = "Number of Days", 
    title = "Heatwave Days", 
    subtitle = "Cape Hatteras, 1971-2021"
  ) +
  theme_minimal(base_family = "Times") +
  scale_x_continuous(breaks = seq(1971, 2021, by = 10)) + 
  scale_fill_manual(
    name = "Measurement",
    breaks = c("count_hse_max", "count_hse_min"),
    labels = c("Max HI", "Min HI"),
    values = c("red", "blue")
  )
```

## WWA GRAPHS
```{r asheville wwa data check}
# load Asheville WWA data
# based on this we see that there were no WWAs issued for Asheville in the last 50 years

avl_wwa = read_excel("data/kavl.wwa.xlsx", col_names = TRUE)
avl_wwa = avl_wwa |>
  select(issued, expired, name)
```

```{r load rdu wwa}
# load RDU WWA data
rdu_wwa = read_excel("data/krdu.wwa.xlsx", col_names = TRUE)
rdu_wwa$year = as.numeric(substr(rdu_wwa$issued, 1, 4)) # year col
rdu_wwa$month = as.numeric(substr(rdu_wwa$issued, 6,7)) # month col
```

```{r wwa counts rdu just summer months}
# wwa counts, rdu, JUST SUMMER MONTHS
rdu_wwa = rdu_wwa |>
  filter(name == "Heat Advisory" | name == "Excessive Heat Watch" | name == "Excessive Heat Warning") |>
  filter(month > 2 & month < 11) |> # filter for mar-oct
  select(year, name) |>
  group_by(year) |>
  summarise(wwa_count = n(),
            .groups = "drop") |>
  subset(year < 2022) |>
  add_row(year = 2013:2014, wwa_count = 0, .before = 9)
```

```{r load clt wwa}
clt_wwa <- read_excel("data/kclt.wwa.xlsx", col_names = TRUE)
clt_wwa$year = as.numeric(substr(clt_wwa$issued, 1, 4))
clt_wwa$month = as.numeric(substr(clt_wwa$issued, 6,7))

clt_wwa = clt_wwa |>
  filter(name == "Heat Advisory" | name == "Excessive Heat Watch" | name == "Excessive Heat Warning") |>
  filter(month > 2 & month < 11) |>
  select(year, name) |>
  group_by(year) |>
  summarise(wwa_count = n(),
            .groups = "drop") |>
  subset(year < 2022) |>
  add_row(year = 2009, wwa_count = 0, .before = 5) |>
  add_row(year = 2013:2015, wwa_count = 0, .before = 9) |>
  add_row(year = 2017:2018, wwa_count = 0, .before = 13) |>
  add_row(year = 2021, wwa_count = 0)
```

```{r load wilm wwa}
wilm_wwa <- read_excel("data/kilm.wwa.xlsx", col_names = TRUE)
wilm_wwa$year = as.numeric(substr(wilm_wwa$issued, 1, 4))
wilm_wwa$month = as.numeric(substr(wilm_wwa$issued, 6,7))

wilm_wwa = wilm_wwa |>
  filter(name == "Heat Advisory" | name == "Excessive Heat Watch" | name == "Excessive Heat Warning") |>
  filter(month > 2 & month < 11) |>
  select(year, name) |>
  group_by(year) |>
  summarise(wwa_count = n(),
            .groups = "drop") |>
  subset(year < 2022) |>
  add_row(year = 2009, wwa_count = 0, .before = 5) |>
  add_row(year = 2014, wwa_count = 0, .before = 10)
```

```{r load hse wwa}
hse_wwa = read_excel("data/khse.wwa.xlsx", col_names = TRUE)
hse_wwa$year = as.numeric(substr(hse_wwa$issued, 1, 4))
hse_wwa$month = as.numeric(substr(hse_wwa$issued, 6,7))

hse_wwa = hse_wwa |>
  filter(name == "Heat Advisory" | name == "Excessive Heat Watch" | name == "Excessive Heat Warning") |>
  filter(month > 2 & month < 11) |>
  select(year, name) |>
  group_by(year) |>
  summarise(wwa_count = n(),
            .groups = "drop") |>
  subset(year < 2022) |>
  add_row(year = 2009, wwa_count = 0, .before = 5) |>
  add_row(year = 2013:2014, wwa_count = 0, .before = 9) |>
  add_row(year = 2021, wwa_count = 0)
```

```{r load gso wwa}
gso_wwa = read_excel("data/kgso.wwa.xlsx", col_names = TRUE)
gso_wwa$year = as.numeric(substr(gso_wwa$issued, 1, 4))
gso_wwa$month = as.numeric(substr(gso_wwa$issued, 6,7))

gso_wwa = gso_wwa |>
  filter(name == "Heat Advisory" | name == "Excessive Heat Watch" | name == "Excessive Heat Warning") |>
  filter(month > 2 & month < 11) |>
  select(year, name) |>
  group_by(year) |>
  summarise(wwa_count = n(),
            .groups = "drop") |>
  subset(year < 2022) |>
  add_row(year = 2008:2009, wwa_count = 0, .before = 4) |>
  add_row(year = 2013:2015, wwa_count = 0, .before = 9)|>
  add_row(year = 2018, wwa_count = 0, .before = 14) |>
  add_row(year = 2021, wwa_count = 0)
```

```{r combine totals from each station into one df}
## combine totals from each station for each year
year_wwa_totals = bind_rows(
  mutate(rdu_wwa, source = "rdu_wwa"),
  mutate(clt_wwa, source = "clt_wwa"),
  mutate(wilm_wwa, source = "wilm_wwa"),
  mutate(hse_wwa, source = "hse_wwa"),
  mutate(gso_wwa, source = "gso_wwa")
)
```

```{r get year counts for each station}
year_wwa_totals = year_wwa_totals |>
  group_by(year) |>
  summarize(total_wwa = sum(wwa_count))
```

```{r filter heatwave days for 2005 and after}
# filter hw days for years 2005 and on
hw_filt = hw |>
  filter(year > 2004)
```

```{r get total number of heatwave days both max and min hi yearly since 2005}
hw_totals_station = hw_filt |>
  mutate(
    avl = count_avl_max + count_avl_min,
    clt = count_clt_max + count_clt_min,
    gso = count_gso_max + count_gso_min,
    hse = count_hse_max + count_hse_min,
    kilm = count_kilm_max + count_kilm_min,
    rdu = count_rdu_max + count_rdu_min
  ) |>
  select(
    c(year, avl, clt, gso, hse, kilm, rdu)
  )
```

```{r get total number of hw days for all stations combined yearly since 2005}
hw_totals = hw_totals_station |>
  mutate(
    hwtotal = avl + clt + gso + hse + kilm + rdu
  ) |>
  select(
    c(year, hwtotal)
    )
```

```{r df with hw and wwa counts per year}
# this is the dataframe containing hw counts and wwa counts
extremes_wwas = data.frame(
  year = hw_totals$year,
  count = c(hw_totals$hwtotal, year_wwa_totals$total_wwa),
  measurement = c(rep("extreme_days", 17), rep("total_wwa", 17))
  )
```

### Plots
```{r bar graph hw days and wwas issued yearly all stations}
extremes_wwas |>
  ggplot(aes(x = year, y = count, fill = measurement)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    x = "Year",
    y = "Count",
    title = "Heatwave Days & Number of WWAs Issued",
    subtitle = "All Stations, 2005-2021"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 4)) +
  theme_minimal(base_family = "Times") +
  scale_fill_manual(
    name = "Measurement",
    breaks = c("extreme_days", "total_wwa"),
    labels = c("Heatwave Days", "WWAs Issued"),
    values = c("#D22B2B", "orange")
  ) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

```{r total number of hw days per year avl}
# individual hw days for avl
hw_totals_avl = hw_totals_station |>
  mutate(
    hwtotal = avl
  ) |>
  select(
    c(year, hwtotal)
    )
```

```{r total number of hw days per year clt}
hw_totals_clt = hw_totals_station |>
  mutate(
    hwtotal = clt
  ) |>
  select(
    c(year, hwtotal)
    )
```

```{r total number of hw days and wwas per year clt}
# get format similar to extremes_wwas
extremes_wwas_clt = data.frame(
  year = hw_totals_clt$year,
  count = c(hw_totals_clt$hwtotal, clt_wwa$wwa_count),
  measurement = c(rep("hw_days", 17), rep("wwa_count", 17))
  )
```

```{r bar graph hw days and wwas issued yearly clt}
extremes_wwas_clt |>
  ggplot(aes(x = year, y = count, fill = measurement)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    x = "Year",
    y = "Count",
    title = "Charlotte"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 4)) +
  scale_y_continuous(breaks = seq(0, 20, by = 5), limits = c(0, 20)) +
  theme_minimal(base_family = "Times") +
  scale_fill_manual(
    name = "Measurement",
    breaks = c("hw_days", "wwa_count"),
    labels = c("Heatwave Days", "WWAs Issued"),
    values = c("#D22B2B", "orange")
  ) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

```{r total number of hw days per year gso}
hw_totals_gso = hw_totals_station |>
  mutate(
    hwtotal = gso
  ) |>
  select(
    c(year, hwtotal)
    )
```

```{r total number of hw days and wwas per year gso}
extremes_wwas_gso = data.frame(
  year = hw_totals_gso$year,
  count = c(hw_totals_gso$hwtotal, gso_wwa$wwa_count),
  measurement = c(rep("hw_days", 17), rep("wwa_count", 17))
  )
```

```{r bar graph hw days and wwas issued yearly gso}
extremes_wwas_gso |>
  ggplot(aes(x = year, y = count, fill = measurement)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    x = "Year",
    y = "Count",
    title = "Greensboro"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 4)) +
  scale_y_continuous(breaks = seq(0, 20, by = 5), limits = c(0, 20)) +
  theme_minimal(base_family = "Times") +
  scale_fill_manual(
    name = "Measurement",
    breaks = c("hw_days", "wwa_count"),
    labels = c("Heatwave Days", "WWAs Issued"),
    values = c("#D22B2B", "orange")
  ) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

```{r total number of hw days per year rdu}
hw_totals_rdu = hw_totals_station |>
  mutate(
    hwtotal = rdu
  ) |>
  select(
    c(year, hwtotal)
    )
```

```{r total number of hw days and wwas per year rdu}
# get format similar to extremes_wwas
extremes_wwas_rdu = data.frame(
  year = hw_totals_rdu$year,
  count = c(hw_totals_rdu$hwtotal, rdu_wwa$wwa_count),
  measurement = c(rep("hw_days", 17), rep("wwa_count", 17))
  )
```

```{r bar graph hw days and wwas issued yearly rdu}
extremes_wwas_rdu |>
  ggplot(aes(x = year, y = count, fill = measurement)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    x = "Year",
    y = "Count",
    title = "Raleigh-Durham"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 4)) +
  scale_y_continuous(breaks = seq(0, 20, by = 5), limits = c(0, 20)) +
  theme_minimal(base_family = "Times") +
  scale_fill_manual(
    name = "Measurement",
    breaks = c("hw_days", "wwa_count"),
    labels = c("Heatwave Days", "WWAs Issued"),
    values = c("#D22B2B", "orange")
  ) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

```{r total number of hw days per year wilm}
hw_totals_wilm = hw_totals_station |>
  mutate(
    hwtotal = kilm
  ) |>
  select(
    c(year, hwtotal)
    )
```

```{r total number of hw days and wwas per year wilm}
# get format similar to extremes_wwas
extremes_wwas_wilm = data.frame(
  year = hw_totals_wilm$year,
  count = c(hw_totals_wilm$hwtotal, wilm_wwa$wwa_count),
  measurement = c(rep("hw_days", 17), rep("wwa_count", 17))
  )
```

```{r bar graph hw days and wwas issued yearly wilm}
extremes_wwas_wilm |>
  ggplot(aes(x = year, y = count, fill = measurement)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    x = "Year",
    y = "Count",
    title = "Wilmington"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 4)) +
  scale_y_continuous(breaks = seq(0, 20, by = 5), limits = c(0, 20)) +
  theme_minimal(base_family = "Times") +
  scale_fill_manual(
    name = "Measurement",
    breaks = c("hw_days", "wwa_count"),
    labels = c("Heatwave Days", "WWAs Issued"),
    values = c("#D22B2B", "orange")
  ) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

```{r total number of hw days per year hse}
hw_totals_hse = hw_totals_station |>
  mutate(
    hwtotal = hse
  ) |>
  select(
    c(year, hwtotal)
    )
```

```{r total number of hw days and wwas per year hse}
# get format similar to extremes_wwas
extremes_wwas_hse = data.frame(
  year = hw_totals_hse$year,
  count = c(hw_totals_hse$hwtotal, hse_wwa$wwa_count),
  measurement = c(rep("hw_days", 17), rep("wwa_count", 17))
  )
```

```{r bar graph hw days and wwas issued yearly hse}
extremes_wwas_hse |>
  ggplot(aes(x = year, y = count, fill = measurement)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    x = "Year",
    y = "Count",
    title = "Cape Hatteras"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 4))+
  scale_y_continuous(breaks = seq(0, 20, by = 5), limits = c(0, 20)) + 
  theme_minimal(base_family = "Times") +
  scale_fill_manual(
    name = "Measurement",
    breaks = c("hw_days", "wwa_count"),
    labels = c("Heatwave Days", "WWAs Issued"),
    values = c("#D22B2B", "orange")
  ) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

the following is for extreme counts, not specifically summer yet
```{r i think just a random years frame}
years = wilm_extreme_counts |>
  filter(year > 2004) |>
  select(year)
```

```{r}
# need some way to plot the same thing but with extreme counts
wilm_filt = wilm_summer_extreme_counts |>
  filter(year > 2004)|>
  select(-c(year))
names(wilm_filt) = c("count_wilm_max", "count_wilm_min")
```

```{r}
rdu_filt = rdu_summer_extreme_counts |>
  filter(year > 2004)|>
  select(-c(year))
names(rdu_filt) = c("count_rdu_max", "count_rdu_min")
```

```{r}
clt_filt = clt_summer_extreme_counts |>
  filter(year > 2004)|>
  select(-c(year))
names(clt_filt) = c("count_clt_max", "count_clt_min")
```

```{r}
gso_filt = gso_summer_extreme_counts |>
  filter(year > 2004)|>
  select(-c(year))
names(gso_filt) = c("count_gso_max", "count_gso_min")
```

```{r}
hse_filt = hse_summer_extreme_counts |>
  filter(year > 2004) |>
  select(-c(year)) 
names(hse_filt) = c("count_hse_max", "count_hse_min")
```

```{r}
extreme_days_filt = cbind(years, clt_filt, gso_filt, hse_filt, wilm_filt, rdu_filt)
```

```{r}
extreme_totals_station = extreme_days_filt |>
  mutate(
    clt = count_clt_max + count_clt_min,
    gso = count_gso_max + count_gso_min,
    hse = count_hse_max + count_hse_min,
    kilm = count_wilm_max + count_wilm_min,
    rdu = count_rdu_max + count_rdu_min
  ) |>
  select(
    c(year, clt, gso, hse, kilm, rdu)
  )
```

```{r}
extreme_totals = extreme_totals_station |>
  mutate(
    extreme_total = clt + gso + hse + kilm + rdu
  ) |>
  select(
    c(year, extreme_total)
    )
```

```{r}
extremesummer_wwas = data.frame(
  year = extreme_totals$year,
  count = c(extreme_totals$extreme_total, year_wwa_totals$total_wwa),
  measurement = c(rep("extreme_days", 17), rep("total_wwa", 17))
  )
```

```{r plot of extreme days and number of wwas issued sliced for summer}
extremesummer_wwas |>
  ggplot(aes(x = year, y = count, fill = measurement)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    x = "Year",
    y = "Count",
    title = "Extreme Days & Number of WWAs Issued",
    subtitle = "All Stations, Mar-Oct, 2005-2021"
  ) +
  scale_x_discrete(breaks = seq(2005, 2021, by = 4)) +
  theme_minimal(base_family = "Times") +
  scale_fill_manual(
    name = "Measurement",
    breaks = c("extreme_days", "total_wwa"),
    labels = c("Extreme Days", "WWAs Issued"),
    values = c("#D22B2B", "orange")
  ) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
extreme_totals_max = extreme_days_filt |>
  mutate(
    clt = count_clt_max,
    gso = count_gso_max,
    hse = count_hse_max,
    kilm = count_wilm_max,
    rdu = count_rdu_max
  ) |>
  select(
    c(year, clt, gso, hse, kilm, rdu)
  )
```

```{r}
extreme_totals_max = extreme_totals_max |>
  mutate(
    extreme_total = clt + gso + hse + kilm + rdu
  ) |>
  select(
    c(year, extreme_total)
    )
```

```{r}
extremesummer_wwas_max = data.frame(
  year = years$year,
  count = c(extreme_totals_max$extreme_total, year_wwa_totals$total_wwa),
  measurement = c(rep("extreme_days", 17), rep("total_wwa", 17))
  )
```

```{r plot extreme days by MAX hi and number of wwas issued sliced for summer}
extremesummer_wwas_max |>
  ggplot(aes(x = year, y = count, fill = measurement)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    x = "Year",
    y = "Count",
    title = "Extreme Days by Daily Max HI & Number of WWAs Issued",
    subtitle = "All Stations, Mar-Oct, 2005-2021"
  ) +
  scale_x_discrete(breaks = seq(2005, 2021, by = 4)) +
  theme_minimal(base_family = "Times") +
  scale_fill_manual(
    name = "Measurement",
    breaks = c("extreme_days", "total_wwa"),
    labels = c("Extreme Days", "WWAs Issued"),
    values = c("#D22B2B", "orange")
  ) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
extreme_totals_min = extreme_days_filt |>
  mutate(
    clt = count_clt_min,
    gso = count_gso_min,
    hse = count_hse_min,
    kilm = count_wilm_min,
    rdu = count_rdu_min
  ) |>
  select(
    c(year, clt, gso, hse, kilm, rdu)
  )
```

```{r}
extreme_totals_min = extreme_totals_min |>
  mutate(
    extreme_total = clt + gso + hse + kilm + rdu
  ) |>
  select(
    c(year, extreme_total)
    )
```

```{r}
extremesummer_wwas_min = data.frame(
  year = years$year,
  count = c(extreme_totals_min$extreme_total, year_wwa_totals$total_wwa),
  measurement = c(rep("extreme_days", 17), rep("total_wwa", 17))
  )
```

```{r plot extreme days by MIN hi and number of wwas issued sliced for summer}
extremesummer_wwas_min |>
  ggplot(aes(x = year, y = count, fill = measurement)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    x = "Year",
    y = "Count",
    title = "Extreme Days by Daily Min HI & Number of WWAs Issued",
    subtitle = "All Stations, Mar-Oct, 2005-2021"
  ) +
  scale_x_discrete(breaks = seq(2005, 2021, by = 4)) +
  theme_minimal(base_family = "Times") +
  scale_fill_manual(
    name = "Measurement",
    breaks = c("extreme_days", "total_wwa"),
    labels = c("Extreme Days", "WWAs Issued"),
    values = c("#D22B2B", "orange")
  ) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

```{r}
clt_diff_hw = hw_totals_clt
clt_diff_hw$wwa_count = clt_wwa$wwa_count

clt_diff_hw = clt_diff_hw |>
  mutate(diff = wwa_count - hwtotal)
```

```{r difference plot clt hw}
clt_diff_hw |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff), levels = c("1", "0", "-1"))))  +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Charlotte"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-15, 15, by = 5)) +  
  scale_fill_manual(values = c("darkgreen", "gray", "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-15, 15))+ 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
gso_diff_hw = hw_totals_gso
gso_diff_hw$wwa_count = gso_wwa$wwa_count

gso_diff_hw = gso_diff_hw |>
  mutate(diff = wwa_count - hwtotal)
```

```{r difference plot gso hw}
gso_diff_hw |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff), levels = c("1", "0", "-1"))))  +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Greensboro"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-15, 15, by = 5)) +  
  scale_fill_manual(values = c("darkgreen", "gray", "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-15, 15))+ 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
rdu_diff_hw = hw_totals_rdu
rdu_diff_hw$wwa_count = rdu_wwa$wwa_count

rdu_diff_hw = rdu_diff_hw |>
  mutate(diff = wwa_count - hwtotal)
```

```{r difference plot rdu hw}
rdu_diff_hw |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff), levels = c("1", "0", "-1")))) +
  geom_bar(stat = "identity", width = 0.6)  +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Raleigh-Durham"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-15, 15, by = 5)) +  
  scale_fill_manual(values = c("darkgreen", "gray", "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-15, 15))+ 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
wilm_diff_hw = hw_totals_wilm
wilm_diff_hw$wwa_count = wilm_wwa$wwa_count

wilm_diff_hw = wilm_diff_hw |>
  mutate(diff = wwa_count - hwtotal)
```

```{r difference plot wilm hw}
wilm_diff_hw |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff), levels = c("1", "0", "-1")))) +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Wilmington"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-15, 15, by = 5)) +  
  scale_fill_manual(values = c("darkgreen", "gray", "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-15, 15))+ 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
hse_diff_hw = hw_totals_hse
hse_diff_hw$wwa_count = hse_wwa$wwa_count

hse_diff_hw = hse_diff_hw |>
  mutate(diff = wwa_count - hwtotal)
```

```{r difference plot hse hw}
hse_diff_hw |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff), levels = c("1", "0", "-1"))))  +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Cape Hatteras"
  ) +
  scale_x_continuous(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-15, 15, by = 5)) +  
  scale_fill_manual(values = c("darkgreen", "gray", "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-15, 15))+ 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
extreme_clt_filt = clt_summer_extreme_counts |>
  filter(year > 2004) |>
  mutate(extreme_count = 
           count_extreme_max + count_extreme_min) |>
  select(-c(count_extreme_max, count_extreme_min))
```

```{r}
clt_diff_extreme = extreme_clt_filt
clt_diff_extreme$wwa_count = clt_wwa$wwa_count

clt_diff_extreme = clt_diff_extreme|>
  mutate(diff = wwa_count - extreme_count)
```

```{r diff plot clt extreme summer}
clt_diff_extreme |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff))))  +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Charlotte"
  ) +
  scale_x_discrete(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-32, 32, by = 8)) +  
  scale_fill_manual(values = c("1" = "darkgreen", "0" = "gray", "-1" = "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-32, 32)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5)
```

```{r}
extreme_gso_filt = gso_summer_extreme_counts |>
  filter(year > 2004) |>
  mutate(extreme_count = 
           count_extreme_max + count_extreme_min) |>
  select(-c(count_extreme_max, count_extreme_min))
```

```{r}
gso_diff_extreme = extreme_gso_filt
gso_diff_extreme$wwa_count = gso_wwa$wwa_count

gso_diff_extreme = gso_diff_extreme|>
  mutate(diff = wwa_count - extreme_count)
```

```{r diff plot gso extreme summer}
gso_diff_extreme |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff))))  +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Greensboro"
  ) +
  scale_x_discrete(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-32, 32, by = 8)) +  
  scale_fill_manual(values = c("1" = "darkgreen", "0" = "gray", "-1" = "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-32, 32)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5)
```

```{r}
extreme_rdu_filt = rdu_summer_extreme_counts |>
  filter(year > 2004) |>
  mutate(extreme_count = 
           count_extreme_max + count_extreme_min) |>
  select(-c(count_extreme_max, count_extreme_min))
```

```{r}
rdu_diff_extreme = extreme_rdu_filt
rdu_diff_extreme$wwa_count = rdu_wwa$wwa_count

rdu_diff_extreme = rdu_diff_extreme|>
  mutate(diff = wwa_count - extreme_count)
```

```{r diff plot rdu extreme summer}
rdu_diff_extreme |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff))))  +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Raleigh-Durham"
  ) +
  scale_x_discrete(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-32, 32, by = 8)) +  
  scale_fill_manual(values = c("1" = "darkgreen", "0" = "gray", "-1" = "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-32, 32)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5)
```

```{r}
extreme_wilm_filt = wilm_summer_extreme_counts |>
  filter(year > 2004) |>
  mutate(extreme_count = 
           count_extreme_max + count_extreme_min) |>
  select(-c(count_extreme_max, count_extreme_min))
```

```{r}
wilm_diff_extreme = extreme_wilm_filt
wilm_diff_extreme$wwa_count = wilm_wwa$wwa_count

wilm_diff_extreme = wilm_diff_extreme|>
  mutate(diff = wwa_count - extreme_count)
```

```{r diff plot wilm extreme summer}
wilm_diff_extreme |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff))))  +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Wilmington"
  ) +
  scale_x_discrete(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-32, 32, by = 8)) +  
  scale_fill_manual(values = c("1" = "darkgreen", "0" = "gray", "-1" = "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-32, 32)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5)
```

```{r}
extreme_hse_filt = hse_summer_extreme_counts |>
  filter(year > 2004) |>
  mutate(extreme_count = 
           count_extreme_max + count_extreme_min) |>
  select(-c(count_extreme_max, count_extreme_min))
```

```{r}
hse_diff_extreme = extreme_hse_filt
hse_diff_extreme$wwa_count = hse_wwa$wwa_count

hse_diff_extreme = hse_diff_extreme|>
  mutate(diff = wwa_count - extreme_count)
```

```{r diff plot hse extreme summer}
hse_diff_extreme |>
  ggplot(aes(x = year, y = diff, fill = factor(sign(diff))))  +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    x = "Year",
    y = "Difference",
    title = "Cape Hatteras"
  ) +
  scale_x_discrete(breaks = seq(2005, 2021, by = 1)) +
  scale_y_continuous(breaks = seq(-32, 32, by = 8)) +  
  scale_fill_manual(values = c("1" = "darkgreen", "0" = "gray", "-1" = "#D22B2B")) +
  theme_bw(base_family = "Times") + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(-32, 32)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_hline(yintercept = 0, color = "black", linetype = "solid", size = 0.5)
```

### Dumbbell Plots
```{r}
clt_diff_extreme |>
  ggplot() + 
  geom_segment(aes(x = year, xend = year, 
                   y = extreme_count, yend = wwa_count), 
               linewidth = 0.8) + 
  geom_point(aes(x = year, y = extreme_count), size = 3, color = "red") + 
  geom_point(aes(x = year, y = wwa_count), size = 3) +
labs(
    x = "Year", 
    y = "Count", 
    title = "Charlotte"
  )  +
  scale_x_discrete(breaks = seq(2005, 2021, by = 2)) + 
  scale_y_continuous(breaks = seq(0, 40, by = 10), limits = c(0, 40)) + 
  theme_classic()  +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
rdu_diff_extreme |>
  ggplot() + 
  geom_segment(aes(x = year, xend = year, 
                   y = extreme_count, yend = wwa_count), 
               linewidth = 0.8) + 
  geom_point(aes(x = year, y = extreme_count), size = 3, color = "red") + 
  geom_point(aes(x = year, y = wwa_count), size = 3) +
labs(
    x = "Year", 
    y = "Count", 
    title = "Raleigh-Durham"
  )  +
  scale_x_discrete(breaks = seq(2005, 2021, by = 2)) + 
  scale_y_continuous(breaks = seq(0, 40, by = 10), limits = c(0, 40)) + 
  theme_classic()  +
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
hse_diff_extreme |>
  ggplot() + 
  geom_segment(aes(x = year, xend = year, 
                   y = extreme_count, yend = wwa_count), 
               linewidth = 0.8) + 
  geom_point(aes(x = year, y = extreme_count), size = 3, color = "red") + 
  geom_point(aes(x = year, y = wwa_count), size = 3) +
labs(
    x = "Year", 
    y = "Count", 
    title = "Cape Hatteras"
  )  +
  scale_x_discrete(breaks = seq(2005, 2021, by = 2)) + 
  scale_y_continuous(breaks = seq(0, 40, by = 10), limits = c(0, 40)) +  
  theme_classic()  +
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
gso_diff_extreme |>
  ggplot() + 
  geom_segment(aes(x = year, xend = year, 
                   y = extreme_count, yend = wwa_count), 
               linewidth = 0.8) + 
  geom_point(aes(x = year, y = extreme_count), size = 3, color = "red") + 
  geom_point(aes(x = year, y = wwa_count), size = 3) +
labs(
    x = "Year", 
    y = "Count", 
    title = "Greensboro"
  )  +
  scale_x_discrete(breaks = seq(2005, 2021, by = 2)) + 
  scale_y_continuous(breaks = seq(0, 40, by = 10), limits = c(0, 40)) +  
  theme_classic()  +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
wilm_diff_extreme |>
  ggplot() + 
  geom_segment(aes(x = year, xend = year, 
                   y = extreme_count, yend = wwa_count), 
               linewidth = 0.8) + 
  geom_point(aes(x = year, y = extreme_count), size = 3, color = "red") + 
  geom_point(aes(x = year, y = wwa_count), size = 3) +
labs(
    x = "Year", 
    y = "Count", 
    title = "Wilmington"
  )  +
  scale_x_discrete(breaks = seq(2005, 2021, by = 2)) + 
  scale_y_continuous(breaks = seq(0, 40, by = 10), limits = c(0, 40)) + 
  theme_classic()  +
  theme(plot.title = element_text(hjust = 0.5))
```


## Unique
```{r}
# load RDU WWA data
rdu_wwa_unique = read_excel("data/krdu.wwa.xlsx", col_names = TRUE)
rdu_wwa_unique$year = as.numeric(substr(rdu_wwa_unique$issued, 1, 4)) # year col
rdu_wwa_unique$month = as.numeric(substr(rdu_wwa_unique$issued, 6,7)) # month col
rdu_wwa_unique$day = as.numeric(substr(rdu_wwa_unique$issued, 9,10)) # day col
```

```{r}
# wwa counts, rdu, JUST SUMMER MONTHS
rdu_wwa_unique = rdu_wwa_unique |>
    filter(name == "Heat Advisory" | name == "Excessive Heat Watch" | name == "Excessive Heat Warning") |>
  filter(month > 2 & month < 11) |> # filter for mar-oct
  select(year, month, day) |>
  group_by(year, month, day) |>
  summarise(daily_count = n(),
            .groups = "drop") 

```

```{r}
rdu_wwa_unique = rdu_wwa_unique |>
  group_by(year) |>
  summarise(wwa_count = n(), 
            .groups = "drop") |>
  subset(year < 2022) |>
  add_row(year = 2013:2014, wwa_count = 0, .before = 9)
```

